
#-------------------- E N V I R O N M E N T -------------------------------------------
DESC: defines functions to be used in pipeline
```{r, echo = FALSE}

rm(list = ls())
#define packages to be used
pkg = c("dplyr", "lubridate", "purrr", "stringr", "ggplot2", "tidyr")
cat("\n Loading required packages: ", pkg, "\n")
pacman::p_load(pkg, character.only = TRUE)

#USER DEFINED FUNCTIONS
# utils <- list.files(file.path(here::here(),"utils"), pattern = "*", full.names = TRUE)
# walk(utils, source)

#SOURCE USED FUNCTIONS
source("R/palettes.R")
source("R/fun_read_list.R")
source("R/key_rename.R")
source("R/fun_wnv_s_clean.R")
source("R/fun_parse_flexible_date.R")
source("R/fun_filter_culex_sheet.R")
source("R/fun_plot_n_trap.R")
source("R/fun_plot_n.R")
source("R/fun_reformat_database_2_culex_sheet.R")
source("R/fun_bind_culex_sheets.R")
source("R/fun_plot_bind_culex.R")       # plot_culex_datasheet_comp
source("R/fun_join_culex_sheets.R")     # join_merge_z
source("R/fun_row_wise_expand.R")       # rowwise_expand
source("R/fun_rm_grp_NA.R")            # rm_grp_na
```

#--------------------------------  C O N F I G  ------------------------------------------------
#DESC: defines variables for functions

```{r}
cfg <- config::get(file = here::here("config", "config_culex_combine.yml"),
                   config = "default") #change to culex if you want to use the culex datasheets

input = file.path(here::here(),cfg$path)
```

#--------------------------------  C O M P I L E  ------------------------------------------------
DESCRIPTION: combines sheets that contain config\$pattern using function read_list

DEPENDENCIES
-   CRAN
    -   dplyr
    -   rio
    -   purrr
    -   tibble
-   utils
    -   fun_read_list
INPUTS
-   config
    -   pattern: pattern for culex files to use in read_list
    
```{r}
culex0 = read_list(path = input, 
                   pattern = cfg$pattern)
```

#--------------------------------  C L E A N  ------------------------------------------------
DESCRIPTION: cleans combined sheets using wnv_s_clean and parse_flexible_date
DEPENDENCIES
-   CRAN
    -   dplyr
    -   lubridate
    -   purrr
    -   ggplot2
-   utils
    -   fun_wnv_s_clean
    -   fun_parse_flexible_date
INPUTS
-   df: dataframe from culex sheets

```{r,clean}
keychain = read.csv(cfg$fn_rename_col)

culex_clean = culex0 %>% 
  key_rename(keychain, drop_extra = TRUE) %>%
  wnv_s_clean()

write.csv(culex_clean, cfg$output_trap_all, row.names = FALSE)

#checked to ensure that data0$trap_type == method and it held true for str_detect(trap_id, "gr) == "G

``` 

#--------------------------------  F I L T E R  ------------------------------------------------
DESCRIPTION: filters cleaned_culex_sheet by a na_col and from a list of traps that you want to keep and removes non culex species
DEPENDENCIES
-   CRAN
    -   dplyr
    -   lubridate
    -   purrr
    -   ggplot2
-   utils
    -   filter_culex_sheet
INPUTS
-   df: dataframe output from wnv_s_clean
-   cfg$fn_trap_keep: filename for the dataframe that contains the traps that you want to keep
-   trap_keep_df: dataframe that contains a column called trap_id that you want to keep from from the culex_sheets

```{r}
# F I L T E R
cat("\n Reading in active foco traps to remove untested traps from", cfg$fn_trap_keep, "\n")

#if filter_active is true will only keep traps from active surveillance tested traps
if(cfg$filter_active){
 trap_keep_df = read.csv(cfg$fn_trap_keep) %>% 
  filter(active == 1)
} else {
  trap_keep_df = read.csv(cfg$fn_trap_keep)
}

temp <- culex_clean |>
  distinct(trap_id, year, week, trap_status, zone, zone2, method)

t = culex_clean %>%
  complete(trap_id, year, week, spp) %>%
  filter(week > 21) %>%
  group_by(trap_id, year, week) %>%
  mutate(
    trap_status = case_when(
      all(is.na(trap_status)) ~ "no mosquitoes",
      any(trap_status == "no culex") ~ "no culex",
      TRUE ~ trap_status
                           )) 
   
culex_filtered = filter_culex_sheet(df = culex_clean, 
                                  na_col = rlang::sym(cfg$na_col), 
                                  trap_keep_df = trap_keep_df)


plot_n_trap(culex_filtered)

skimr::skim(culex_filtered)

write.csv(culex_filtered, cfg$output_culex, row.names = FALSE)
```

#------------------------ R E F O R M A T   D A T A S H E E T -------------------------------------
DESCRIPTION: we only have culex sheets from 2019 to 2024 and CDC didn't send culex sheets so we need to reformat
the wnv-s_database that has pooled data to get the missing data

DEPENDENCIES
-   CRAN
    -   dplyr
-   utils
    -   reformat_database
INPUTS
-   cfg$fn_database: filename for the csv that has the wnv-s_database

```{r}
database = read.csv(cfg$fn_database)

database_reformat = database %>%
  reformat_database_2_culex_sheet() %>%
  wnv_s_clean(rm_col = "trap_status")
```

#------------------------ B I N D   D A T A S H E E T  &  C U L E X -------------------------------------
DESCRIPTION: bind the culex data with the wnv-s_database that has been reformatted to match the culex sheet
in order to visualize the difference in mosquito numbers between the culex sheet (trap data) and the
database (pool data)

DEPENDENCIES
-   CRAN
    -   dplyr
    -   ggplot2
-   utils
    -   bind_culex_sheets
    -   plot_culex_datasheet_comp
INPUTS
-   culex_filtered:culex sheet that only has the active traps in surveillance
-   data_reformat: the wnv-s_database that has been reformatted

```{r}
# Restrict database to the same active traps as culex_filtered so the
# comparison is apples-to-apples. Without this filter, inactive LV/FC traps
# in the database inflate the datasheet totals above the culex sheet totals.
culex_database_bind = bind_culex_sheets(
  culex_filtered,
  database_reformat %>% dplyr::semi_join(trap_keep_df, by = "trap_id")
)

# Filter to the first year with culex-sheet data so years with only database
# records don't appear as "pools > traps" (which is scientifically impossible).
culex_start_yr <- min(culex_filtered$year, na.rm = TRUE) - 1

p_comp = plot_culex_datasheet_comp(culex_database_bind %>% filter(!zone %in% cfg$filter_zone),
                                   culex_start_yr)
p_comp

ggsave(plot = p_comp, filename = cfg$output_plot_comp)
```

#------------------------ J O I N   D A T A S H E E T  &  C U L E X -------------------------------------
DESCRIPTION: join the culex data with the wnv-s_database that has been reformatted to match the culex sheet
in order to visualize the difference in mosquito numbers between the culex sheet (trap data) and the
database (pool data)

DEPENDENCIES
-   CRAN
    -   dplyr
    -   ggplot2
-   utils
    -   join_merge_z
    -   plot_culex_datasheet_comp
INPUTS
-   culex_filtered:culex sheet that only has the active traps in surveillance
-   data_reformat: the wnv-s_database that has been reformatted


```{r}
# Join on (trap_id, zone, year, week, spp) only.
# database_reformat's spp0 = "Tarsalis"/"Pipiens" (abbreviated) while culex_filtered's
# spp0 = "Culex tarsalis"/"Culex pipiens" (full name). Including spp0 — or trap_date,
# zone2, method — in the join key prevents any rows from matching. Trimming to the
# minimal key columns forces a clean merge on the biologically meaningful key.
database_reformat_trim = database_reformat %>%
  dplyr::select(trap_id, zone, year, week, spp, total)

culex_database = join_merge_z(culex_filtered, database_reformat_trim, "total")

write.csv(culex_database$full, cfg$output_culex_database)

```




#------------------------ A D D   N A  C U L E X   T R A P  -------------------------------------
DESCRIPTION: using the foco_trap get the NA traps from the culex sheet to test the completeness of those sheets

DEPENDENCIES
-   CRAN
    -   dplyr
    -   ggplot2
-   utils
    -   bind_culex_sheets
INPUTS
-   data_filtered:
-   data_reformat: the wnv-s_database that has been reformatted

```{r, add na}
y = lubridate::year(Sys.Date())

if(cfg$filter_active){
 trap_keep_df = read.csv(cfg$fn_trap_keep) %>% 
  filter(active == 1)
} else {
  trap_keep_df = read.csv(cfg$fn_trap_keep)
}

trap_keep_df = trap_keep_df %>%
  mutate(end = y,
         w_start = 23,
         w_end = 37)

trap_expand = rowwise_expand(trap_keep_df, "start", "end", "year")
trap_expand = rowwise_expand(trap_expand, "w_start", "w_end", "week")
trap_expand = trap_expand %>%
  select(trap_id, zone, year, week)

culex_database_expand = left_join(trap_expand, culex_database$final, 
                    by = colnames(trap_expand))

#remove years and trap ids where we have no data which indicates the traps didn't exist yet or we don't have any data
culex_database_expand = culex_database_expand %>%
  #rm_grp_na(c("year", "week", "zone"), "total") %>%
  mutate(total = replace_na(total, 0)) %>%
  mutate(spp = replace_na(spp, "No Mosquitoes")) %>%
  # Step 1: propagate dates from any trap in the same zone-week (fills NAs for unobserved weeks)
  group_by(year, week, zone) %>%
  fill(trap_date, .direction = "downup") %>%
  ungroup() %>%
  # Step 2: normalize to one date per trap-week so wnv_s_clean groups all rows together.
  # Without this, culex "no mosquitoes" rows (set date) and database rows (collection date)
  # land in separate group_by(trap_id, trap_date) groups → same trap gets two trap_status
  # values in the same week → inflates the plot count above the expected number of traps.
  group_by(trap_id, zone, year, week) %>%
  mutate(trap_date = dplyr::first(sort(trap_date, na.last = TRUE))) %>%
  ungroup() %>%
  wnv_s_clean()


t = culex_database_expand %>% 
  dplyr::mutate(zone = dplyr::if_else(zone %in% cfg$fc_zone, 
                                    "FC", 
                                    zone)) %>%
    filter(is.null(zone)) %>%
    group_by(year, week, zone) %>%
    summarise(n = n_distinct(trap_id),
              .groups = "drop") %>%
  filter(zone == "FC" & n != 53)

#t2 = culex_database_expand %>% filter(zone %in% cfg$fc_zone & year == 2020 & week == 23)


p_trap_n = plot_n_trap(culex_database_expand) +
  ggtitle("N traps by Zone")

p_trap_n
ggsave(plot = p_trap_n, filename = "3_output/trap_status.png")
```


#------------------------ G E T   N A  C U L E X   T R A P  -------------------------------------
DESCRIPTION: using the foco_trap get the NA traps from the culex sheet to test the completeness of those sheets

DEPENDENCIES
-   CRAN
    -   dplyr
    -   ggplot2
-   utils
    -   bind_culex_sheets
INPUTS
-   data_filtered:
-   data_reformat: the wnv-s_database that has been reformatted

```{r}
y = lubridate::year(Sys.Date())-1
BE_rm = c("LC-061", "LC-058", "LC-062", "LC-059", "LC-060") #remove the traps cdc sent us and weren't managed by VDCI at the time

trap_keep_df = trap_keep_df %>%
  filter(active == 1) %>%
  filter(!trap_id %in% BE_rm) %>%
  mutate(end = y,
         w_start = 23,
         w_end = 37)

trap_expand = rowwise_expand(trap_keep_df, "start", "end", "year")
trap_expand = rowwise_expand(trap_expand, "w_start", "w_end", "week")
trap_expand = trap_expand %>%
  mutate(week = week,
         year = year) %>%
  select(trap_id, zone, year, week)

na_trap = anti_join(trap_expand, culex_clean, 
                    by = c("trap_id", "year", "week")) %>%
  mutate(zone = if_else(zone %in% cfg$fc_zone, "FC", zone)) %>%
  filter(year > 2018) %>%
  filter(zone %in% c("FC", "LV"))

p_na_trap = plot_n(na_trap, trap_id) +
  ggtitle("Number of weeks trap is missing. 2019-2024")

p_na_trap

ggsave(plot = p_na_trap, filename = "3_output/na_trap.png")
```



