
#-------------------- E N V I R O N M E N T -------------------------------------------
DESC: defines functions to be used in pipeline
```{r, echo = FALSE}

rm(list = ls())
#define packages to be used
pkg = c("dplyr", "lubridate", "purrr", "stringr", "ggplot2", "tidyr")
cat("\n Loading required packages: ", pkg, "\n")
pacman::p_load(pkg, character.only = TRUE)

#USER DEFINED FUNCTIONS
# utils <- list.files(file.path(here::here(),"utils"), pattern = "*", full.names = TRUE)
# walk(utils, source)

#SOURCE USED FUNCTIONS
source("R/palettes.R")
source("R/fun_read_list.R")
source("R/key_rename.R")
source("R/fun_wnv_s_clean.R")
source("R/fun_parse_flexible_date.R")
source("R/fun_filter_culex_sheet.R")
source("R/fun_plot_n_trap.R")
source("R/fun_plot_n.R")
source("R/fun_reformat_database_2_culex_sheet.R")
source("R/fun_bind_culex_sheets.R")
source("R/fun_plot_bind_culex.R")       # plot_culex_datasheet_comp
source("R/fun_join_culex_sheets.R")     # join_merge_z
source("R/fun_row_wise_expand.R")       # rowwise_expand
source("R/fun_rm_grp_NA.R")            # rm_grp_na
source("R/fun_fix_zone.R")             # fix_zone_from_ref
source("R/fun_plot_all_spp.R")
source("R/fun_dedup_culex.R")       # culex_dedup
source("R/fun_make_key.R")
source("R/fun_expand_trap_spp.R")
```

#--------------------------------  C O N F I G  ------------------------------------------------
#DESC: defines variables for functions

```{r}
cfg <- config::get(file = here::here("config", "config_culex_combine.yml"),
                   config = "default") #change to culex if you want to use the culex datasheets

input = file.path(here::here(),cfg$path)
```

#---------------------------  I M P O R T  ------------------------------------------------
```{r}
trap_ref = read.csv(cfg$fn_trap_keep) #foco_trap
keychain = read.csv(cfg$fn_rename_col) #column renaming key
database = read.csv(cfg$fn_database) #databaes for augmenting culex sheets
```


#--------------------------- B U L K   I M P O R T --------------------------------------------
DESCRIPTION: combines sheets that contain config\$pattern using function read_list

DEPENDENCIES
-   CRAN
    -   dplyr
    -   rio
    -   purrr
    -   tibble
-   utils
    -   fun_read_list
INPUTS
-   config
    -   pattern: pattern for culex files to use in read_list
    
```{r}
culex0 = read_list(path = input, 
                   pattern = cfg$pattern)

write.csv(culex0, cfg$mid_culex0, row.names = FALSE)
```



#------------------------  E X P A N D   G R I D  ------------------------------------------------
```{r}
#data = expand_grid(trap_ref)

```



#--------------------------------  C L E A N  ------------------------------------------------
DESCRIPTION: cleans combined sheets using wnv_s_clean and parse_flexible_date
DEPENDENCIES
-   CRAN
    -   dplyr
    -   lubridate
    -   purrr
    -   ggplot2
-   utils
    -   fun_wnv_s_clean
    -   fun_parse_flexible_date
INPUTS
-   df: dataframe from culex sheets

```{r,clean}

culex_clean = culex0 %>%
  key_rename(keychain, drop_extra = TRUE) %>%
  wnv_s_clean() %>%
  make_key(key_cols = c("trap_id", "spp", "year", "week")) %>%
  culex_dedup()   # remove weekend collection dates (non-standard surveillance nights)


culex_clean_dupes_key = culex_clean %>% get_dupes(key) %>% filter(spp != "non culex")
culex_clean_dupes_all = culex_clean %>% get_dupes() %>% filter(spp != "non culex")

write.csv(culex_clean, cfg$output_trap_all, row.names = FALSE)

p_culex_clean = plot_all_spp(culex_clean, "zone2")
p_culex_clean
ggsave(cfg$mid_culex_clean_plot, p_culex_clean, width = 12, height = 10)

#checked to ensure that data0$trap_type == method and it held true for str_detect(trap_id, "gr) == "G

``` 

#--------------------------------  F I L T E R  ------------------------------------------------
DESCRIPTION: filters cleaned_culex_sheet by a na_col and from a list of traps that you want to keep and removes non culex species
DEPENDENCIES
-   CRAN
    -   dplyr
    -   lubridate
    -   purrr
    -   ggplot2
-   utils
    -   filter_culex_sheet
INPUTS
-   df: dataframe output from wnv_s_clean
-   cfg$fn_trap_keep: filename for the dataframe that contains the traps that you want to keep
-   trap_keep_df: dataframe that contains a column called trap_id that you want to keep from from the culex_sheets

```{r, filter}

# F I L T E R
cat("\n Reading in active foco traps to remove untested traps from", cfg$fn_trap_keep, "\n")

#if filter_active is true will only keep traps from active surveillance tested traps
if(cfg$filter_active){
 trap_keep_df = read.csv(cfg$fn_trap_keep) %>% 
  filter(active == 1)
} else {
  trap_keep_df = read.csv(cfg$fn_trap_keep)
}

#WARNING FOR BERTHOUD TRAPS EXPECTED 
culex_filtered = filter_culex_sheet(df = culex_clean, 
                                  na_col = rlang::sym(cfg$na_col), 
                                  trap_keep_df = trap_keep_df)


plot_n_trap(culex_filtered)

skimr::skim(culex_filtered)

write.csv(culex_filtered, cfg$mid_culex_database, row.names = FALSE)
```

#------------------------ R E F O R M A T   D A T A S H E E T -------------------------------------
DESCRIPTION: we only have culex sheets from 2019 to 2024 and CDC didn't send culex sheets so we need to reformat
the wnv-s_database that has pooled data to get the missing data

DEPENDENCIES
-   CRAN
    -   dplyr
-   utils
    -   reformat_database
INPUTS
-   cfg$fn_database: filename for the csv that has the wnv-s_database

```{r}


# Load the full (unfiltered) trap reference so inactive traps also get correct zones.
# This is separate from trap_keep_df which may be filtered to active == 1.

database_reformat = database %>%
  reformat_database_2_culex_sheet() %>%
  fix_zone_from_ref(trap_ref) %>%    # correct zone from foco_trap - data.csv before zone2 is derived
  wnv_s_clean() %>%
  make_key(key_cols = c("trap_id", "spp", "year", "week")) %>%
  culex_dedup()   # remove weekend collection dates (non-standard surveillance nights)


database_reformat_dupes_key = database_reformat %>% get_dupes(key)
database_reformat_dupes_all = database_reformat %>% get_dupes()

write.csv(database_reformat, cfg$mid_database_reformat, row.names = FALSE)
```

#------------------------ P L O T   D A T A S H E E T  vs  C U L E X -------------------------------------
DESCRIPTION: bind the culex data with the wnv-s_database that has been reformatted to match the culex sheet
in order to visualize the difference in mosquito numbers between the culex sheet (trap data) and the
database (pool data)

DEPENDENCIES
-   CRAN
    -   dplyr
    -   ggplot2
-   utils
    -   bind_culex_sheets
    -   plot_culex_datasheet_comp
INPUTS
-   culex_filtered:culex sheet that only has the active traps in surveillance
-   data_reformat: the wnv-s_database that has been reformatted

```{r}
# Restrict database to the same active traps as culex_filtered so the
# comparison is apples-to-apples. Without this filter, inactive LV/FC traps
# in the database inflate the datasheet totals above the culex sheet totals.
culex_database_bind = bind_culex_sheets(
  culex_filtered,
  database_reformat %>% dplyr::semi_join(trap_keep_df, by = "trap_id")
)

# Filter to the first year with culex-sheet data so years with only database
# records don't appear as "pools > traps" (which is scientifically impossible).
culex_start_yr <- min(culex_filtered$year, na.rm = TRUE) - 1

p_comp = plot_culex_datasheet_comp(culex_database_bind %>% filter(!zone %in% cfg$filter_zone),
                                   culex_start_yr)
p_comp

ggsave(plot = p_comp, filename = cfg$output_plot_comp)
```

#------------------------ J O I N   D A T A S H E E T  &  C U L E X -------------------------------------
DESCRIPTION: join the culex data with the wnv-s_database that has been reformatted to match the culex sheet
in order to visualize the difference in mosquito numbers between the culex sheet (trap data) and the
database (pool data)

DEPENDENCIES
-   CRAN
    -   dplyr
    -   ggplot2
-   utils
    -   join_merge_z
    -   plot_culex_datasheet_comp
INPUTS
-   culex_filtered:culex sheet that only has the active traps in surveillance
-   data_reformat: the wnv-s_database that has been reformatted


```{r}
# join_merge_z joins on (trap_id, spp, year, week) and coalesces all shared
# columns, preferring culex_sheet values. The database fills in trap-weeks that
# are absent from the culex sheet entirely (e.g., CDC Berthoud traps in 2024
# before VDCI took over the contract).
#
# source is added here so the coalesce loop in join_merge_z resolves it:
#   matched rows   -> "culex_sheet"  (.x wins)
#   culex-only     -> "culex_sheet"
#   database-only  -> "database"
culex_database = join_cxsheet_database(
  culex_filtered    %>% dplyr::mutate(source = "culex_sheet"),
  database_reformat %>% dplyr::mutate(source = "database")
)

culex_database_dupes_key = culex_database %>% get_dupes(key)
culex_database_dupes_all= culex_database %>% get_dupes()

write.csv(culex_database, cfg$output_culex_database, row.names = FALSE)

```




#------------------------ A D D   N A  C U L E X   T R A P  -------------------------------------
DESCRIPTION: using the foco_trap get the NA traps from the culex sheet to test the completeness of those sheets

DEPENDENCIES
-   CRAN
    -   dplyr
    -   ggplot2
-   utils
    -   bind_culex_sheets
INPUTS
-   data_filtered:
-   data_reformat: the wnv-s_database that has been reformatted

```{r, add na}
# Load and filter the active trap reference list (cfg$filter_active from config)
if (cfg$filter_active) {
  trap_keep_df <- read.csv(cfg$fn_trap_keep) %>%
    dplyr::filter(active == 1)
} else {
  trap_keep_df <- read.csv(cfg$fn_trap_keep)
}

# expand_trap_spp() builds trap × year × week × spp grid, left-joins to
# culex_database, fills missing spp with total = 0, and flags zone-weeks
# with no data at all as trap_status = "No Traps" (trap_date = NA).
# wnv_s_clean() is called outside the function to standardise column types
# and derive zone2, method, and trap_status from the expanded output.
culex_database_expand <- expand_trap_spp(
  trap_keep_df   = trap_keep_df,
  culex_database = culex_database,
  spp_levels     = cfg$spp_levels
) %>%
  wnv_s_clean()


write.csv(culex_database_expand, cfg$output_culex_database_expand, row.names = FALSE)

p_trap_n = plot_n_trap(culex_database_expand) +
  ggtitle("N traps by Zone")

p_trap_n
ggsave(plot = p_trap_n, filename = "3_output/trap_status.png", width = 12, height = 6)

```


#------------------------ G E T   N A  C U L E X   T R A P  -------------------------------------
DESCRIPTION: using the foco_trap get the NA traps from the culex sheet to test the completeness of those sheets

DEPENDENCIES
-   CRAN
    -   dplyr
    -   ggplot2
-   utils
    -   bind_culex_sheets
INPUTS
-   data_filtered:
-   data_reformat: the wnv-s_database that has been reformatted

```{r}
y = lubridate::year(Sys.Date())-1

trap_keep_df = trap_keep_df %>%
  filter(active == 1) %>%
  mutate(end = y,
         w_start = 23,
         w_end = 37)

trap_expand = rowwise_expand(trap_keep_df, "start", "end", "year")
trap_expand = rowwise_expand(trap_expand, "w_start", "w_end", "week")
trap_expand = trap_expand %>%
  mutate(week = week,
         year = year) %>%
  select(trap_id, zone, year, week)

na_trap = anti_join(trap_expand, culex_clean, 
                    by = c("trap_id", "year", "week")) %>%
  mutate(zone = if_else(zone %in% cfg$fc_zone, "FC", zone)) %>%
  filter(year > 2018) %>%
  filter(zone %in% c("FC", "LV"))

p_na_trap = plot_n(na_trap, trap_id) +
  ggtitle("Number of weeks trap is missing. 2019-2024")

p_na_trap

ggsave(plot = p_na_trap, filename = "3_output/na_trap.png")
```



