
#-------------------- E N V I R O N M E N T -------------------------------------------
DESC: defines functions to be used in pipeline
```{r, echo = FALSE}
# Set working directory to project root so relative source() paths resolve correctly.
knitr::opts_knit$set(root.dir = here::here())

rm(list = ls())
#define packages to be used
pkg = c("dplyr", "lubridate", "purrr", "stringr", "ggplot2", "tidyr", "janitor")
cat("\n Loading required packages: ", pkg, "\n")
pacman::p_load(pkg, character.only = TRUE)

#USER DEFINED FUNCTIONS
# utils <- list.files(file.path(here::here(),"utils"), pattern = "*", full.names = TRUE)
# walk(utils, source)

#SOURCE USED FUNCTIONS
source("R/palettes.R")
source("R/fun_read_list.R")
source("R/fun_read_source.R")
source("R/fun_pivot_cmc.R")
source("R/fun_pivot_cfc.R")
source("R/fun_pivot_boulder.R")
source("R/fun_manifest.R")
source("R/key_rename.R")
source("R/fun_wnv_s_clean.R")
source("R/fun_parse_flexible_date.R")
source("R/fun_filter_culex_sheet.R")
source("R/fun_plot_n_trap.R")
source("R/fun_plot_n.R")
source("R/fun_reformat_database_2_culex_sheet.R")
source("R/fun_bind_culex_sheets.R")
source("R/fun_plot_bind_culex.R")       # plot_culex_datasheet_comp
source("R/fun_join_culex_sheets.R")     # join_merge_z
source("R/fun_row_wise_expand.R")       # rowwise_expand
source("R/fun_rm_grp_NA.R")            # rm_grp_na
source("R/fun_fix_zone.R")             # fix_zone_from_ref
source("R/fun_plot_all_spp.R")
source("R/fun_dedup_culex.R")       # culex_dedup
source("R/fun_make_key.R")
source("R/fun_expand_trap.R")          # expand_trap
source("R/fun_assign_trap_status.R")   # assign_trap_status_post_expand
source("R/fun_fill_skeleton.R")        # fill_skeleton, collect_unmerged
source("R/fun_prep_for_skeleton.R")    # prep_for_skeleton
```

#--------------------------------  C O N F I G  ------------------------------------------------
#DESC: defines variables for functions

```{r}
cfg <- config::get(file = here::here("config", "config_culex_combine.yml"),
                   config = "default")

input = file.path(here::here(), cfg$path)
cfg$manifest <- here::here(cfg$manifest)  # resolve to absolute so wd changes don't matter

manifest_init(cfg$manifest)
manifest_log(cfg$manifest, "config",
             config_file = "config/config_culex_combine.yml",
             year_start  = cfg$year_start,
             week_range  = paste(cfg$week_start, "-", cfg$week_end),
             spp_levels  = cfg$spp_levels)
```

#---------------------------  I M P O R T  ------------------------------------------------
```{r}
trap_ref = read.csv(cfg$fn_trap_keep) #foco_trap
keychain = read.csv(cfg$fn_rename_col) #column renaming key
database = read.csv(cfg$fn_database) #databaes for augmenting culex sheets
```


#---------------------------  P E R   S O U R C E   I M P O R T  ------------------------------
# DESC: read each source declared in cfg$sources, dispatch the optional pivot_fn,
# then funnel every source through the canonical clean chain.

```{r, per_source_import}
# YAML stores pivot function names as strings (e.g. pivot_fn: "pivot_cmc").
# pivot_registry translates those strings to the actual R function objects
# so they can be passed to read_source(). Sources with pivot_fn: null get NULL.
pivot_registry <- list(
  pivot_cmc     = pivot_cmc,    # wide CMC sheet  → long
  pivot_cfc     = pivot_cfc,    # CFC column rename only (already long)
  pivot_boulder = pivot_boulder  # Boulder pass-through, enforces zone = "BC"
)

# imap iterates cfg$sources (a named list from config_culex_combine.yml).
#   src_name = the source key, e.g. "cmc_weekly"
#   src      = the config sub-list for that source (path, pattern, pivot_fn, …)
#
# For each source:
#   1. Resolve the pivot_fn string → function object via pivot_registry
#      e.g. src$pivot_fn == "pivot_cmc"  →  pivot_fn = pivot_cmc
#   2. Call read_source(), which lists files, drops exclude_paths, reads each
#      file with rio::import (+ any import_args like sheet="CombinedData"),
#      calls pivot_fn(raw_df) if not NULL, coerces all columns to character,
#      and binds rows into one tibble.
#
# Result: sources_raw — a named list with one raw tibble per source key.
sources_raw <- purrr::imap(cfg$sources, function(src, src_name) {
  # Look up the pivot function by name; NULL means the file is already long.
  pivot_fn      <- if (is.null(src$pivot_fn))      NULL        else pivot_registry[[src$pivot_fn]]
  exclude_paths <- if (is.null(src$exclude_paths)) character() else src$exclude_paths
  import_args   <- if (is.null(src$import_args))   list()      else src$import_args
  read_source(
    path          = src$path,
    pattern       = src$pattern,
    pivot_fn      = pivot_fn,
    exclude_paths = exclude_paths,
    import_args   = import_args
  )
})

# Inventory files per source for the manifest
purrr::iwalk(cfg$sources, function(src, src_name) {
  inv <- inventory_source(src$path, src$pattern)
  manifest_log(cfg$manifest, paste0("inventory:", src_name),
               path    = src$path,
               pattern = src$pattern,
               n_files = inv$n_files)
})

# Raw row counts per source
purrr::iwalk(sources_raw, function(df, src_name) {
  cat(sprintf("raw %-12s rows: %d\n", src_name, nrow(df)))
  manifest_log(cfg$manifest, paste0("read:", src_name),
               n_rows = nrow(df), n_cols = ncol(df))
})
```



#------------------------  E X P A N D   G R I D  ------------------------------------------------
```{r}
# Build skeleton FIRST — the complete trap×year×week×spp grid that all
# observed data is joined into. No cleaning happens here.
trap_skeleton <- expand_trap(
  trap_ref   = trap_ref,
  spp_levels = cfg$spp_levels,
  year_start = cfg$year_start,
  year_end   = lubridate::year(Sys.Date()),
  week_start = cfg$week_start,
  week_end   = cfg$week_end,
  fc_zones   = cfg$fc_zone
)

cat("Skeleton rows:", nrow(trap_skeleton), "\n")
cat("Skeleton traps:", dplyr::n_distinct(trap_skeleton$trap_id), "\n")
```



#---------------------------  P E R   S O U R C E   C L E A N  --------------------------------
# DESC: clean_one() does ONLY cleaning — column normalisation, date/zone/
# species/status assignment, and dedup. It does NOT drop rows for analysis
# reasons (no spp filter, no active-trap filter). The output retains every
# observed species (including non-culex) and every observed trap.
#
# prep_for_skeleton() and filter_culex_sheet() now live in their own
# downstream chunks (per_source_prep and per_source_filter) so each stage
# has its own row-count log in the manifest.

```{r, per_source_clean}
# clean_one applies the canonical cleaning chain to one source's raw tibble.
# Arguments:
#   df_raw   — raw tibble from sources_raw (all columns are character at this point)
#   src_name — source key string, used only for manifest logging
#
# Chain (clean only — no filtering, no skeleton prep):
#   key_rename()  → translates source-specific column names to canonical names
#                    (trap_id, trap_date, spp, total, …) using database_column_rename.csv
#   wnv_s_clean() → parses dates, derives zone2/year/week/method, classifies spp,
#                    sets trap_status (culex / malfunction / no_mosquitoes / no culex);
#                    preserves the original spp value in column spp0
#   make_key()    → builds a composite key "trap_id|spp|year|week" used for dedup
#   culex_dedup() → drops non-standard collection nights (weekends, duplicate pickups)
#
# Returns: tibble with all observed species and all observed traps. spp0
# carries the raw species label for use by plot_all_spp().
clean_one <- function(df_raw, src_name) {
  if (nrow(df_raw) == 0L) {
    manifest_log(cfg$manifest, paste0("clean:", src_name), status = "EMPTY — skipped")
    return(tibble::tibble(
      key = character(), trap_id = character(), zone = character(),
      zone2 = character(), year = integer(), week = integer(),
      trap_date = as.Date(character()), method = character(),
      spp = character(), spp0 = character(),
      trap_status = character(), total = numeric()
    ))
  } #end of nrow == 0

  after_clean <- df_raw %>%
    key_rename(keychain, drop_extra = TRUE) %>%
    wnv_s_clean()
  after_dedup <- after_clean %>%
    make_key(key_cols = c("trap_id", "spp", "year", "week")) %>%
    culex_dedup()

  manifest_log(cfg$manifest, paste0("clean:", src_name),
               n_raw            = nrow(df_raw),
               n_after_clean    = nrow(after_clean),
               n_after_dedup    = nrow(after_dedup),
               removed_by_dedup = nrow(after_clean) - nrow(after_dedup))

  after_dedup
} #end clean_one

# imap iterates sources_raw (same names as cfg$sources):
#   .x = the raw tibble for this source
#   .y = the source name string (e.g. "cmc_weekly")
# Result: sources_cleaned — a named list, one cleaned tibble per source.
sources_cleaned <- purrr::imap(sources_raw, ~ clean_one(.x, .y))
purrr::iwalk(sources_cleaned, ~ cat(sprintf("cleaned %-12s rows: %d\n", .y, nrow(.x))))

# All-species combined view for the diagnostic plot and the all-spp CSV.
# spp0 is preserved by wnv_s_clean() and holds the original species label
# (e.g. "Aedes vexans", "Culex erythrothorax", "non culex") — plot_all_spp
# groups by spp0 so this view shows every species observed, not just the
# Tarsalis/Pipiens subset that ends up in the skeleton.
culex_clean <- dplyr::bind_rows(sources_cleaned, .id = "source")
write.csv(culex_clean, cfg$output_trap_all, row.names = FALSE)

p_culex_clean <- plot_all_spp(culex_clean, "zone2")
p_culex_clean
ggsave(cfg$mid_culex_clean_plot, p_culex_clean, width = 12, height = 10)
```

#---------------------------  P E R   S O U R C E   P R E P  ----------------------------------
# DESC: prep_for_skeleton() expands each cleaned tibble so every row has
# spp ∈ cfg$spp_levels. Status events (malfunction / no culex / no mosquitoes)
# get expanded to one row per spp_level with total set to NA or 0. Non-target
# species (Aedes, "non culex", etc.) are dropped here — they are kept upstream
# in sources_cleaned for the all-spp diagnostic plot.

```{r, per_source_prep}
sources_prepped <- purrr::imap(sources_cleaned, function(df, src_name) {
  if (nrow(df) == 0L) {
    manifest_log(cfg$manifest, paste0("prep:", src_name), status = "EMPTY — skipped")
    return(df)
  }
  out <- prep_for_skeleton(df, cfg$spp_levels)
  manifest_log(cfg$manifest, paste0("prep:", src_name),
               n_in_clean    = nrow(df),
               n_out_prepped = nrow(out))
  out
})
purrr::iwalk(sources_prepped, ~ cat(sprintf("prepped %-12s rows: %d\n", .y, nrow(.x))))
```

#--------------------------  P E R   S O U R C E   F I L T E R  -------------------------------
# DESC: apply filter_culex_sheet() per source based on each source's
# filter_active flag in config_culex_combine.yml. Sources with filter_active:
# false (e.g. Boulder) pass through unchanged. The active-trap reference is
# loaded here and reused by the comparison plot downstream.

```{r, per_source_filter}
if (isTRUE(cfg$filter_active)) {
  trap_keep_df <- read.csv(cfg$fn_trap_keep) %>% dplyr::filter(active == 1)
} else {
  trap_keep_df <- read.csv(cfg$fn_trap_keep)
}

sources_filtered <- purrr::imap(sources_prepped, function(df, src_name) {
  if (nrow(df) == 0L) {
    manifest_log(cfg$manifest, paste0("filter:", src_name), status = "EMPTY — skipped")
    return(df)
  }
  src <- cfg$sources[[src_name]]
  out <- if (isTRUE(src$filter_active)) {
    filter_culex_sheet(df,
                       na_col       = rlang::sym(cfg$na_col),
                       trap_keep_df = trap_keep_df)
  } else {
    df
  }
  manifest_log(cfg$manifest, paste0("filter:", src_name),
               filter_active     = isTRUE(src$filter_active),
               n_in_prepped      = nrow(df),
               n_out_filtered    = nrow(out),
               removed_by_filter = nrow(df) - nrow(out))
  out
})
purrr::iwalk(sources_filtered, ~ cat(sprintf("filtered %-12s rows: %d\n", .y, nrow(.x))))

# Capture rows dropped by the active-trap semi_join inside filter_culex_sheet().
# Without this sink, trap_ids that are NOT in foco_trap (e.g. typo'd names like
# "FC-033075") disappear silently — they never reach fill_skeleton(), so they
# never appear in unmerged_data.csv. The mirror logic below applies the same
# NA-clean + distinct steps that filter_culex_sheet() performs before its
# semi_join, then anti_joins trap_keep_df to capture exactly the rows removed
# by the active-trap filter (case 3 of filter_culex_sheet's three drops).
dropped_by_filter <- purrr::imap_dfr(sources_prepped, function(df, src_name) {
  src <- cfg$sources[[src_name]]
  if (!isTRUE(src$filter_active) || nrow(df) == 0L) return(tibble::tibble())
  df %>%
    dplyr::filter(!is.na(.data[[cfg$na_col]]), nchar(.data[[cfg$na_col]]) > 0) %>%
    dplyr::distinct() %>%
    dplyr::anti_join(trap_keep_df, by = cfg$na_col) %>%
    dplyr::mutate(source = src_name)
})

write.csv(dropped_by_filter, cfg$mid_dropped_by_filter, row.names = FALSE)

manifest_log(cfg$manifest, "filter:dropped_by_active_filter",
             n_total_dropped   = nrow(dropped_by_filter),
             n_unique_trap_ids = dplyr::n_distinct(dropped_by_filter$trap_id),
             output_path       = cfg$mid_dropped_by_filter)
```

#--------------------------------  F I L T E R  ------------------------------------------------
DESCRIPTION: bind the per-source filtered tibbles into a single dataframe for
downstream plots, the database join, and the skeleton fill.
DEPENDENCIES
-   CRAN
    -   dplyr
    -   lubridate
    -   purrr
    -   ggplot2
-   utils
    -   filter_culex_sheet (applied in per_source_filter chunk above)
INPUTS
-   sources_filtered: named list produced by per_source_filter chunk

```{r, filter}
# culex_filtered = combined post-filter sources for downstream plots and validations.
# Per-source filtering was applied in the per_source_filter chunk above.
culex_filtered <- dplyr::bind_rows(
  purrr::keep(sources_filtered, function(df) nrow(df) > 0)
)

plot_n_trap(culex_filtered)

skimr::skim(culex_filtered)

write.csv(culex_filtered, cfg$mid_culex_database, row.names = FALSE)
```

#------------------------ R E F O R M A T   D A T A S H E E T -------------------------------------
DESCRIPTION: we only have culex sheets from 2019 to 2024 and CDC didn't send culex sheets so we need to reformat
the wnv-s_database that has pooled data to get the missing data

DEPENDENCIES
-   CRAN
    -   dplyr
-   utils
    -   reformat_database
INPUTS
-   cfg$fn_database: filename for the csv that has the wnv-s_database

```{r}


# Load the full (unfiltered) trap reference so inactive traps also get correct zones.
# This is separate from trap_keep_df which may be filtered to active == 1.

database_reformat = database %>%
  reformat_database_2_culex_sheet() %>%
  fix_zone_from_ref(trap_ref) %>%    # correct zone from foco_trap - data.csv before zone2 is derived
  wnv_s_clean() %>%
  make_key(key_cols = c("trap_id", "spp", "year", "week")) %>%
  culex_dedup()   # remove weekend collection dates (non-standard surveillance nights)


database_reformat_dupes_key = database_reformat %>% get_dupes(key)
database_reformat_dupes_all = database_reformat %>% get_dupes()

write.csv(database_reformat, cfg$mid_database_reformat, row.names = FALSE)

database_reformat_prepped <- prep_for_skeleton(database_reformat, cfg$spp_levels)
```

#------------------------ P L O T   D A T A S H E E T  vs  C U L E X -------------------------------------
DESCRIPTION: bind the culex data with the wnv-s_database that has been reformatted to match the culex sheet
in order to visualize the difference in mosquito numbers between the culex sheet (trap data) and the
database (pool data)

DEPENDENCIES
-   CRAN
    -   dplyr
    -   ggplot2
-   utils
    -   bind_culex_sheets
    -   plot_culex_datasheet_comp
INPUTS
-   culex_filtered:culex sheet that only has the active traps in surveillance
-   data_reformat: the wnv-s_database that has been reformatted

```{r}
# Restrict database to the same active traps as culex_filtered so the
# comparison is apples-to-apples. Without this filter, inactive LV/FC traps
# in the database inflate the datasheet totals above the culex sheet totals.
culex_database_bind = bind_culex_sheets(
  culex_filtered,
  database_reformat %>% dplyr::semi_join(trap_keep_df, by = "trap_id")
)

# Filter to the first year with culex-sheet data so years with only database
# records don't appear as "pools > traps" (which is scientifically impossible).
culex_start_yr <- min(culex_filtered$year, na.rm = TRUE) - 1

p_comp = plot_culex_datasheet_comp(culex_database_bind %>% filter(!zone %in% cfg$filter_zone),
                                   culex_start_yr)
p_comp

ggsave(plot = p_comp, filename = cfg$output_plot_comp)
```

#------------------------ J O I N   D A T A S H E E T  &  C U L E X -------------------------------------
DESCRIPTION: join the culex data with the wnv-s_database that has been reformatted to match the culex sheet
in order to visualize the difference in mosquito numbers between the culex sheet (trap data) and the
database (pool data)

DEPENDENCIES
-   CRAN
    -   dplyr
    -   ggplot2
-   utils
    -   join_merge_z
    -   plot_culex_datasheet_comp
INPUTS
-   culex_filtered:culex sheet that only has the active traps in surveillance
-   data_reformat: the wnv-s_database that has been reformatted


```{r}
# join_merge_z joins on (trap_id, spp, year, week) and coalesces all shared
# columns, preferring culex_sheet values. The database fills in trap-weeks that
# are absent from the culex sheet entirely (e.g., CDC Berthoud traps in 2024
# before VDCI took over the contract).
#
# source is added here so the coalesce loop in join_merge_z resolves it:
#   matched rows   -> "culex_sheet"  (.x wins)
#   culex-only     -> "culex_sheet"
#   database-only  -> "database"
# NOTE: culex_database is used for the comparison plot below (culex_v_datasheet).
# The skeleton join uses culex_filtered and database_reformat_prepped directly so that
# each source enters the skeleton independently rather than through a pre-merged
# dataset — preserving the skeleton's role as the deduplication boundary.
culex_database = join_cxsheet_database(
  culex_filtered    %>% dplyr::mutate(source = "culex_sheet"),
  database_reformat %>% dplyr::mutate(source = "database")
)

culex_database_dupes_key = culex_database %>% get_dupes(key)
culex_database_dupes_all= culex_database %>% get_dupes()

write.csv(culex_database, cfg$output_culex_database, row.names = FALSE)

```


#------------------------ J O I N   I N T O   S K E L E T O N  ----------------------------------
# DESCRIPTION: fill the skeleton from each source in priority order using
# fill_skeleton(). The culex sheet is preferred — it fills as many cells as
# possible first. The database patches only cells still NA after the culex join.
# Adding a new data source (CMC, Boulder, etc.) is a single list entry here.

# zone2 and method come from the skeleton (expand_trap() from trap_ref) and are
# never overwritten by source data.

# collect_unmerged() captures records from any source absent from the skeleton
# (traps observed but not in the active reference list). Saved for audit.
```{r, join_skeleton}
# Priority order: VDCI -> CMC weekly -> CFC -> BC -> database.
# Each list name becomes the value of the `source` column in the output.
#
# TODO (see docs/TODO.md): when the separate all-mosquito pipeline lands,
# it will emit a small handoff CSV — one row per (trap_id, year, week) where
# all_mosq observed activity but no culex were caught. Read it here as a
# LOW-priority source and append to data_sources BEFORE the database entry,
# e.g.:
#   trap_was_set <- read.csv("2_mid/trap_was_set.csv") %>%
#     prep_for_skeleton(cfg$spp_levels)
#   data_sources <- c(
#     purrr::keep(sources_filtered, ~ nrow(.x) > 0L),
#     list(trap_was_set = trap_was_set,
#          database     = database_reformat_prepped)
#   )
# Effect: skeleton rows that were active traps but caught only non-culex
# (currently land in unmerged or get labelled "no trap" by
# assign_trap_status_post_expand) will be correctly labelled "no culex" —
# fixing the abundance-denominator bias for 2020–2024.
data_sources <- c(
  purrr::keep(sources_filtered, ~ nrow(.x) > 0L),
  list(database = database_reformat_prepped)
)

culex_database_expand <- fill_skeleton(trap_skeleton, data_sources)

unmerged <- collect_unmerged(trap_skeleton, data_sources)
write.csv(unmerged, cfg$mid_unmerged_data, row.names = FALSE)

purrr::iwalk(split(unmerged, unmerged$source),
             ~ cat(sprintf("Unmerged %-12s rows: %d\n", .y, nrow(.x))))

# Per-source join accounting
purrr::iwalk(sources_filtered, function(df, src_name) {
  if (nrow(df) == 0L) return()
  n_in  <- nrow(df)
  n_unm <- sum(unmerged$source == src_name, na.rm = TRUE)
  n_exp <- sum(culex_database_expand$source == src_name, na.rm = TRUE)
  manifest_log(cfg$manifest, paste0("join:", src_name),
               n_into_skeleton   = n_in,
               n_in_final_expand = n_exp,
               n_unmerged        = n_unm,
               balance_check     = paste0(n_in, " == ", n_exp, " + ", n_unm,
                                          " : ",
                                          if (n_in == n_exp + n_unm) "OK"
                                          else paste0("DRIFT (", n_in - n_exp - n_unm, ")")))
})
```

#------------------------ P O S T - E X P A N D   S T A T U S  ----------------------------------
# DESCRIPTION: resolve trap_status for skeleton rows with no data match.
# Zone-week present in data → "no mosquitoes"; zone-week entirely absent → "no trap".
```{r, post_expand_status}

culex_database_expand <- assign_trap_status_post_expand(culex_database_expand)

# ── VALIDATIONS ──────────────────────────────────────────────────────────────
# These are hard assertions, not soft warnings. If either fails the pipeline
# output is wrong and must be investigated before use.
#
# Baseline: culex_filtered (active traps, post-filter), NOT culex_clean.
# culex_clean includes inactive traps that are never in the skeleton by design.
# Comparing against culex_clean would count trap-weeks that cannot appear in
# the output, causing false failures. culex_filtered is the correct baseline
# because it has already been restricted to the same active-trap scope as the skeleton.

# 1. Malfunction trap-week count must be identical in input and final output.
#    Malfunctions carry no count data (total = NA); any drop means a trap event
#    was lost entirely from the analysis denominator.
#    Restrict to skeleton-eligible keys: historical sources (CMC, CFC) may have
#    malfunction records for traps/weeks outside the skeleton scope that correctly
#    land in unmerged rather than in culex_database_expand.
n_raw_malfunction <- culex_filtered %>%
  dplyr::filter(trap_status == "malfunction") %>%
  dplyr::semi_join(trap_skeleton, by = "key") %>%
  dplyr::distinct(trap_id, year, week) %>%
  nrow()

n_final_malfunction <- culex_database_expand %>%
  dplyr::filter(trap_status == "malfunction") %>%
  dplyr::distinct(trap_id, year, week) %>%
  nrow()

cat(
  "\nValidation — malfunction trap-weeks:\n",
  "  culex_filtered input: ", n_raw_malfunction, "\n",
  "  final expanded:       ", n_final_malfunction, "\n"
)
stopifnot(
  "Malfunction trap-week count changed between input and final output" =
    n_final_malfunction == n_raw_malfunction
)

# 2. "No mosquitoes" trap-week count must be ≥ culex sheet input count.
#    Expansion adds zero-fill rows so final count should grow, never shrink.
#    Restrict to skeleton-eligible keys so historical out-of-scope records do
#    not inflate the baseline and cause false failures.
n_raw_no_mosq <- culex_filtered %>%
  dplyr::filter(trap_status == "no mosquitoes") %>%
  dplyr::semi_join(trap_skeleton, by = "key") %>%
  dplyr::distinct(trap_id, year, week) %>%
  nrow()

n_final_no_mosq <- culex_database_expand %>%
  dplyr::filter(trap_status == "no mosquitoes") %>%
  dplyr::distinct(trap_id, year, week) %>%
  nrow()

cat(
  "\nValidation — 'no mosquitoes' trap-weeks:\n",
  "  culex_filtered input: ", n_raw_no_mosq, "\n",
  "  final expanded:       ", n_final_no_mosq, "\n"
)
stopifnot(
  "'no mosquitoes' trap-weeks shrank between input and final output" =
    n_final_no_mosq >= n_raw_no_mosq
)

# 3. "No culex" trap-week count must be identical in input and final output.
#    Traps that caught only non-culex species contribute to the denominator
#    and must not be silently reclassified as "no mosquitoes" or "no trap".
#    Baseline is culex_filtered restricted to skeleton-eligible keys via semi_join.
#    Some "no culex" trap-weeks in culex_filtered fall outside the skeleton's
#    year/week scope (23-37) and correctly appear in unmerged — excluding them
#    from the baseline prevents false failures on out-of-season records.
n_raw_no_culex <- culex_filtered %>%
  dplyr::filter(trap_status == "no culex") %>%
  dplyr::semi_join(trap_skeleton, by = "key") %>%
  dplyr::distinct(trap_id, year, week) %>%
  nrow()

n_final_no_culex <- culex_database_expand %>%
  dplyr::filter(trap_status == "no culex") %>%
  dplyr::distinct(trap_id, year, week) %>%
  nrow()

cat(
  "\nValidation — 'no culex' trap-weeks:\n",
  "  culex_filtered input: ", n_raw_no_culex, "\n",
  "  final expanded:       ", n_final_no_culex, "\n"
)
stopifnot(
  "'no culex' trap-week count changed between input and final output" =
    n_final_no_culex == n_raw_no_culex
)

# 4. malfunction and no trap rows must have total = NA.
stopifnot(
  "malfunction and no trap rows must have total = NA" = {
    culex_database_expand %>%
      dplyr::filter(trap_status %in% c("malfunction", "no trap"),
                    !is.na(total)) %>%
      nrow() == 0
  }
)

# 5. no culex and no mosquitoes rows must have total = 0.
stopifnot(
  "no culex and no mosquitoes rows must have total = 0" = {
    culex_database_expand %>%
      dplyr::filter(trap_status %in% c("no culex", "no mosquitoes"),
                    is.na(total) | total != 0) %>%
      nrow() == 0
  }
)

# 6. Completeness — every Pipiens/Tarsalis row from every source's filtered
#    input must appear in either culex_database_expand or unmerged. A drop
#    here means a record was silently lost between filter and join.
all_prepped_keys <- dplyr::bind_rows(sources_filtered) %>%
  dplyr::filter(spp %in% cfg$spp_levels) %>%
  dplyr::pull(key) %>%
  unique()

accounted_keys <- unique(c(
  culex_database_expand$key,
  unmerged$key
))

missing_keys <- setdiff(all_prepped_keys, accounted_keys)

cat(
  "\nValidation — completeness:\n",
  "  prepped keys:   ", length(all_prepped_keys), "\n",
  "  accounted keys: ", length(accounted_keys), "\n",
  "  missing keys:   ", length(missing_keys), "\n"
)
stopifnot(
  "Source prepped records dropped silently between prep and skeleton join" =
    length(missing_keys) == 0L
)

manifest_log(cfg$manifest, "final",
             skeleton_rows        = nrow(trap_skeleton),
             expand_rows          = nrow(culex_database_expand),
             unmerged_rows        = nrow(unmerged),
             malfunction_trap_wks = n_final_malfunction,
             no_mosq_trap_wks     = n_final_no_mosq,
             no_culex_trap_wks    = n_final_no_culex)

write.csv(culex_database_expand, cfg$output_culex_database_expand, row.names = FALSE)
```

#------------------------ P L O T  ----------------------------------
# DESCRIPTION: Plot results for visual checks
# Zone-week present in data → "no mosquitoes"; zone-week entirely absent → "no trap".
```{r, plot}

# spp column: raw species label used for validation alongside trap_status
p_culex_spp <- plot_all_spp(culex_database_expand, "zone2", spp_col = "spp")
p_culex_spp

ggplot2::ggsave(plot = p_culex_spp, filename = "3_output/culex_counts.png",
                width = 12, height = 6)

p_trap_n <- plot_n_trap(culex_database_expand) +
  ggplot2::ggtitle("N traps by Zone")
p_trap_n
ggplot2::ggsave(plot = p_trap_n, filename = "3_output/trap_status.png",
                width = 12, height = 6)
```
